Analysis and Figures for Gallego et al, 2020

### 2020-09-05 Scripts/Manuscript.Main.Analysis.Rmd
Wilcoxon made easy

expanded.future <- future.scenario %>% 
  select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units)

## Get

@@

```{r Fig3left}

expanded.future %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species))    %>%
  mutate(Area = as.character(Area)) %>%
  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                          TRUE          ~ Area)) %>%
  mutate(Year = fct_rev(as.factor(Year) )) -> input.for.left

input.for.left %>%
   group_by(phylum, Temperature, pH, Year, Area ) %>%
  summarise(present = sum(present)) %>%
  group_by(phylum,Area) %>% 
  mutate(present = present/max(present)) %>% 
  mutate (label = phylum) -> phylum.stats

phylum.stats %>% 
  nest(-phylum) %>% 
  mutate( plot.by.phyl = map2 (data, phylum, ~ .x %>%
  ggplot(aes()) +
  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                    breaks=c("2017","2095")) +
  labs (x = "Phylum-wide standardized richness",
        y = "") +
  facet_wrap(~Area) +
  theme_minimal() +
  theme (strip.text.y = element_text(angle = 0),
        axis.text.y = element_text(face = "italic"), 
        #plot.background   = element_rect(fill = "#E5E9F0"),
        legend.direction = "horizontal",
        legend.position  = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold",vjust =  -5 ),
        axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
        geom_histogram(aes( x= present,after_stat(ncount),
                            fill = as.factor(Year),
                            group= as.factor(Year)),
                       binwidth = 0.025,
                       colour = "black",
                       position = "identity", 
                       alpha = 0.75) +
        scale_x_continuous(position = "top", limits = c(0,1)))) %>% 
  select(-data)-> summary.plot

input.for.left %>% 
  semi_join(second.sel, by = "taxa") %>%
  nest(-phylum) %>% 
  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
                                 ggplot(aes(x= present,
                                           y = label)) +
                                 scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                                                   breaks=c("2017","2095")) +
                                 labs (x = "Probability of presence",
                                       y = "") +
                                 facet_wrap(~Area) +
                                 theme_minimal() +
                                 theme(strip.text.y = element_text(angle = 0),
                                       axis.text.y = element_text(face = "italic"),
                                       legend.direction = "horizontal",
                                       legend.position  = "bottom",
                                       plot.title.position = "plot",
                                       plot.title = element_text(face = "bold",vjust =  -5 ),
                                       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
                                 ggtitle(.y)  +
                                 geom_boxplot(aes(fill = as.factor(Year)), outlier.alpha = 0.5) +
                                 scale_x_continuous(position = "top") )) -> fig3.left

@@

full.grid$sd <- apply(posterior.predict.all.taxa,2,sd )

@@

Using a global ANOVA to see if there are differences between Years (2095-2017) in both Areas. Then a  wilcoxon test to see if there is a difference  on each phylum
```{r}

@@

input.for.left %>% 
  mutate(sim.row = rep(1:10873, n_distinct(taxa))) -> input.richness

input.richness %>% 
  nest(-Area) %>% 
  mutate(data = map (data, ~ .x %>% 
                       mutate(present = case_when(present < 0.5 ~ 0,
                                                  TRUE          ~ 1)) %>% 
                         group_by(Year,  sim.row) %>% 
                         summarise( Richness = sum(present)) ),
         wilcoxon = map(data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater")),
         tidy_wilcox = map(wilcoxon, broom::tidy)) %>%
  pull(wilcoxon)

@@

input.richness %>% 
  nest(-phylum,-Area) %>% 
  #  slice(1) %>%
    mutate(data = map (data, ~ .x %>% 
                         mutate(present = case_when(present < 0.5 ~ 0,
                                                  TRUE          ~ 1)) %>% 
                         group_by(Year,  sim.row) %>% 
                         summarise( Richness = sum(present)) )) %>% 
   mutate(wilcoxon = map (data, ~ wilcox.test(Richness ~ Year, data = .x, alternative = "greater") ))  -> results.lm

results.lm %>%  filter (phylum == "Bacillariophyta") %>% pull(wilcoxon) 
```

>>>>

expanded.future <- future.scenario %>% 
  select(Area, Year, Temperature = Temp.std, TemperatureCelsius, pH = pH.std, pH_units) 
nsims <- nrow(future.scenario)
## Get

@@

```{r Fig3left}
expanded.future %>% 
  separate(taxa, into = c("family","genus", "species"), sep = "\\|", remove = F) %>%
  mutate(label = case_when((species == "NA" & genus == "NA") ~ paste0(family, " sp."),
                           species == "NA" ~  paste0(genus, " sp."),
                           TRUE            ~ species))    %>%
  mutate(Area = as.character(Area)) %>%
  mutate(Area = case_when(Area == "SJI" ~ "San Juan Island",
                          TRUE          ~ Area)) %>%
  mutate(Year = fct_rev(as.factor(Year) )) -> input.for.left

input.for.left %>% 
  group_by(taxa) %>% 
  nest %>% 
  mutate(data = map (data, ~.x %>% 
                       rownames_to_column("sim.row"))) %>% 
  unnest -> input.for.left

```{r Fig3left botto}
input.for.left %>%
   group_by(phylum, sim.row,Temperature, pH, Year, Area ) %>%
  summarise(present = sum(present)) %>%
  group_by(phylum,Area) %>% 
  mutate(present = present/max(present)) %>% 
  mutate (label = phylum) -> phylum.stats

phylum.stats %>% 
  nest(-phylum) %>% 
  mutate( plot.by.phyl = map2 (data, phylum, ~ .x %>%
  ggplot(aes()) +
  scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                    breaks=c("2017","2095")) +
  labs (x = "Phylum-wide standardized richness",
        y = "") +
  facet_wrap(~Area) +
  theme_minimal() +
  theme (strip.text.y = element_text(angle = 0),
        axis.text.y = element_text(face = "italic"), 
        #plot.background   = element_rect(fill = "#E5E9F0"),
        legend.direction = "horizontal",
        legend.position  = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(face = "bold",vjust =  -5 ),
        axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
        geom_histogram(aes( x= present,after_stat(ncount),
                            fill = as.factor(Year),
                            group= as.factor(Year)),
                       binwidth = 0.025,
                       colour = "black",
                       position = "identity", 
                       alpha = 0.75) +
        scale_x_continuous(position = "top", limits = c(0,1)))) %>% 
  select(-data)-> summary.plot
```

```{r Fig3left top}
input.for.left %>% 
  semi_join(second.sel, by = "taxa") %>%
  nest(-phylum) %>% 
  mutate( plot.by.taxa = map2 (data, phylum, ~ .x %>%
                                 ggplot(aes(x= present,
                                           y = label)) +
                                 scale_fill_brewer(name = "Year",  type = "qual", palette = 4, direction = 1,
                                                   breaks=c("2017","2095")) +
                                 labs (x = "Probability of presence",
                                       y = "") +
                                 facet_wrap(~Area) +
                                 theme_minimal() +
                                 theme(strip.text.y = element_text(angle = 0),
                                       axis.text.y = element_text(face = "italic"),
                                       legend.direction = "horizontal",
                                       legend.position  = "bottom",
                                       plot.title.position = "plot",
                                       plot.title = element_text(face = "bold",vjust =  -5 ),
                                       axis.title.x.top  = element_text(face = "bold",vjust = 3) ) +
                                 ggtitle(.y)  +
                                 geom_boxplot(aes(fill = as.factor(Year)), outlier.alpha = 0.5) +
                                 scale_x_continuous(position = "top") )) -> fig3.left

@@

@@

First calculate the richness in each point as following. In each environmental point, the 100 draws represent if a taxa was found or not. We sum the draws of all taxa so we have 100 estimation of richness for each point - we then take the mean of those 100 draws and we have a estimation of richness at that point.

```{r get the richness per phylum}
input.for.left %>%
   group_by(phylum, sim.row,Temperature, pH, Year, Area ) %>%
  summarise(richness = sum(present)) -> richness.by.phylum.by.point

# And from here - make the global

richness.by.phylum.by.point %>% 
  group_by (Area,Year, sim.row) %>% 
  summarise(richness = sum(richness)) -> richness.by.point

Now the Global - richness by point

#overall change in richness
richness.by.point %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
#diatom change in richness
richness.by.phylum.by.point %>%
  filter(phylum == "Bacillariophyta") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
```

Because we are comparing two 
```{r}

@@

richness.by.point %>% 
  nest(-Area) %>% 
  mutate(wilcoxon = map(data, ~ wilcox.test(richness ~ Year, data = .x)),
         tidy_wilcox = map_df(wilcoxon, broom::tidy))

@@

richness.by.phylum.by.point %>% 
   nest(-phylum,-Area) %>% 
   mutate(wilcoxon = map (data, ~ wilcox.test(richness ~ Year, data = .x) ),
          wiloc.tidy= map_df(wilcoxon, broom::tidy))  -> results.lm

results.lm %>%  filter (phylum == "Bacillariophyta")
```
